A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-03-18 graph_syntax.tex
updated the Desgin Variable, Objectives, Constraints section with some new notes. Two new node types

\section{MDAO Graph Primatives}
There are three node types:  
\begin{description}
\item[variable node] represents passing of data.
\item[model node] represents the handling of data.
\item[driver node] represents constrol structures capable of managing iteration
\end{description}

Only the variable node and the model node are allowed to be used in an MCG or FPF. 
All three nodes are required for any SPF. IN addition to the three node types, there 
are two edge types:

@@

Via the analysis block structure, we can also distinguish between input and ouptut
variable nodes. Inputs are any variable nodes that have an outgoing edge into a model 
node. Conversely, outputs are any variable nodes that have an incoming edge from a model node. 
Note that by this definition, inputs and outputs can only exist when variable nodes poses an 
edge joining them to a model node. In Fig. \ref{f:analysis block} these are labeled out as 
local inputs and local outputs. The necessity to be connected to a specific model node forces
the variable nodes to be local in nature. When information from the output of one analysis 
block is passed, or connected, to the input of another analysis block, a new free edge is added 
connecting the two variable nodes involved in the exchange. The edge is free because, unlike the edges 
within an analysis block, it could be added or removed depending on the demands of the specific problem.

@@

The MDO problem formulation concepts represented by nodes are given in 
Table \ref{t:node representation}, and the concepts represented by edges are 
given in Table \ref{t:edge representation}.
% \begin{itemize}
% \item An objective or constraint is indicated by a variable node with no outgoing edges and with the incoming edges being directed from only other variable nodes.
% \item A global input is a variable node with no edges directed in and with at least two edges directed out to different variable nodes.
% \item A design variable is represented by a variable node with no edges directed in and on fixed edge directed out.
% \end{itemize}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by a node}
  \label{t:node representation}
  \begin{tabular}{ccc} \hline 
edges directed in & edges directed out & node representation \\ \hline
only free edges & none & objective or constraint\\
none & at least two free edges & global input \\
none & one fixed edge & design variable \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by an edge}
  \label{t:edge representation}
  \begin{tabular}{ccc} \hline 
from node & to node & edge representation \\ \hline
variable & variable & connection/passing of a variable\\
variable & model & local input \\
model & variable & local output \\
model & model & function call \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
%`Connection' edges connect the local outputs from one analysis block to input nodes.

\subsection{Local vs Global Properties}

>>>>

\section{MDAO Graph Primatives}
There are four node types:  
\begin{description}
\item[variable] represents scalar or array data
\item[model] responsible for mapping inputs to outputs
\item[expression] simple algebreic combinations of variables (such as for constraint of objective functions)
\item[driver] constrol structures capable of managing iteration
\end{description}

Only the variable, model nodes are allowed to be used in an MCG. In a FPF expression 
nodes are also required. All four nodes are required for any SPF. In addition to 
the three node types, there are two edge types:

@@

Via the analysis block structure, we can also distinguish between input and ouptut
variable nodes. Inputs are any variable nodes that have an outgoing edge into a model 
node. Conversely, outputs are any variable nodes that have an incoming edge from a model node. 
Note that by this definition, inputs and outputs can only exist when variable nodes poses an 
edge joining them to a model node. In Fig. \ref{f:analysis block} these are labeled out as 
local inputs and local outputs. The variables are local because they are explicitely tied
to the analysis block and have no relevance without it.

MDAO problems require that information be passed between sets of analyses. When information from the output of one analysis 
block is passed, or connected, to the input of another analysis block, a new free edge is added 
connecting the two variable nodes involved in the exchange. The edge is free because, unlike the edges 
within an analysis block, it could be added or removed depending on the demands of the specific problem.

@@

Design variables, objectives, and constraints all share the property, that they are 
require external decisions to be made about a given system to identify them. Consider 
a notional problem given by the graph in Fig. [insert ref]a. There are three input variable 
nodes which, could potentially become design variables: w, x, and y. Assume that the 
output y is connected to the input y, yeilding Fig. [insert ref]b. Now only w and x could 
be design variables. We define a design variable to be any input variable node that does 
not have an incoming connection.

[[INSERT GRAPH HERE]]

Objective and constraints are also constructs that need to be identified by engineers 
as part of the design process. In the case of objective functions single output values 
could be selected, but commonly multiple values are assembeled together via simple 
athrimetic opperations to form the final objective function. For example, the problem from 
Fig. [insert ref] might have an objective function of $z^2$ or $2y + z$. Constraints are 
always given in the form of either an inequality of an equality, e.g. 
$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer explictly identifies variables 
or groups of variables and assigns an simple, explicit relationship between them.

For the proposed syntax objectives and constraints are represented through the use of 
expression nodes. These nodes contain the actual mathematical expression used to define the 
objective or constraint value. You can distinguish between an objective or constraint node 
via the presence of a simple expression (objective) vs an inequality/equality (constraint). 
Fig. [insert new ref] shows the notional problem graph augmented with objective and 
constraint values.

[[INSERT ANOTHER GRAPH HERE]]

\begin{comment}
The MDO problem formulation concepts represented by nodes are given in 
Table \ref{t:node representation}, and the concepts represented by edges are 
given in Table \ref{t:edge representation}.
% \begin{itemize}
% \item An objective or constraint is indicated by a variable node with no outgoing edges and with the incoming edges being directed from only other variable nodes.
% \item A global input is a variable node with no edges directed in and with at least two edges directed out to different variable nodes.
% \item A design variable is represented by a variable node with no edges directed in and on fixed edge directed out.
% \end{itemize}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by a node}
  \label{t:node representation}
  \begin{tabular}{ccc} \hline 
edges directed in & edges directed out & node representation \\ \hline
only free edges & none & objective or constraint\\
none & at least two free edges & global input \\
none & one fixed edge & design variable \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
\begin{table}[h!]
 \begin{center}
  \caption{Problem formulation concept represented by an edge}
  \label{t:edge representation}
  \begin{tabular}{ccc} \hline 
from node & to node & edge representation \\ \hline
variable & variable & connection/passing of a variable\\
variable & model & local input \\
model & variable & local output \\
model & model & function call \\ \hline
  \end{tabular}
 \end{center}
 \vspace{-15pt}
\end{table}
%`Connection' edges connect the local outputs from one analysis block to input nodes.
\end{comment}
\subsection{Local vs Global Properties}

\begin{itemize}
  \item what can be local or global? Variables, objectives, constraints, 
  \item What does it really mean to be local vs global? 
  \item What does the graph structure of local vs global look like? 
  \item local is a frame of reference... a variable could be global when looking at two analyses, but local to the two when looking at the three
\end{itemize}
